---
title: "basic_usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{basic_usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(plinkr)
```

In this vignette, I run the 
commands at [http://zzz.bwh.harvard.edu/plink/data.shtml#plink](http://zzz.bwh.harvard.edu/plink/data.shtml#plink) using plinkr.

```
plink --file mydata --pheno pheno.raw --assoc --maf 0.05 --out run1
```

The `--file mydata` means that PLINK will look for the files named
`mydata.ped` and `mydata.map`. In this example, we will use these two from
the PLINK example files:

```{r}
# Check both files exists
ped_filename <- get_plink_example_filename("test.ped")
map_filename <- get_plink_example_filename("test.map")
# Get the path without extension, can use either ped or map file
testthat::expect_equal(
  tools::file_path_sans_ext(ped_filename),
  tools::file_path_sans_ext(map_filename)
)
# Get the path without extension
mydata <- tools::file_path_sans_ext(ped_filename)
```


The `--pheno pheno.raw` means that PLINK will look for a files named `pheno.raw`. PLINK does not supply this example file, hence we will use one supplied by plinkr:

```{r}
phenotype_filename <- get_plinkr_filename("pheno.raw")
```

becomes

```{r}
plinkr::run_plink(
  args = c(
    "--file", mydata, # mydata.ped and mydata.map
    "--pheno", phenotype_filename, 
    "--assoc", 
    "--maf", "0.05", 
    "--out", "run1"
  )
)
```


Ru-run if (`rerun_plink`) is fun for the future:

```
plink --rerun run1.log --maf 0.1
plink --rerun run1.log --maf 0.1 --out run2
```

Run PLINK from script, not now (`run_plink_from_script`):

```
plink --script myscript1.txt
```

> where the file myscript1.txt is a plain text file containing

```
--ped ..\data\version1\50K\allsamples.ped 
--map ..\data\allmapfiles\finalversion\autosomal.map 
--out ..\results\working\sample-missingness-v1.22
--from rs66537222
--to rs8837323 
--geno 0.25
--maf 0.02
--missing
```

