---
title: "plinkr_example_files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plinkr_example_files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette shows how to plinkr example files were created.

```{r setup}
library(plinkr)
```

# Methods

To demonstrate how the plinkr example files were created,
these are re-created in a temporary folder, 
and then compared to the existing files
in the `plinkr` `inst/extdata` folder

Determine the name of the temporary folder:

```{r get_temp_folder_name}
folder_name <- plinkr::get_plinkr_tempfilename()
folder_name
```

## `demo_assoc`

Simulate data:

```{r}
# Create the example files in the inst/extdata folder
set.seed(314)
assoc_params <- create_demo_assoc_params()
```

Glance at the data:

```{r}
knitr::kable(assoc_params$data$map_table)
```

```{r}
knitr::kable(assoc_params$data$ped_table)
```

Should be the same:

```{r}
testthat::expect_equal(
  assoc_params$data$map_table,
  read_plink_map_file(get_plinkr_filename("demo_assoc.map"))
)
```

```{r}
testthat::expect_equal(
  assoc_params$data$ped_table,
  read_plink_ped_file(get_plinkr_filename("demo_assoc.ped"))
)
```

## `demo_assoc_qt` 

### plain-text files

```{r}
set.seed(314)
assoc_qt_params <- create_demo_assoc_qt_params()
```

Glance at the data:

```{r}
knitr::kable(assoc_qt_params$data$map_table)
```

```{r}
knitr::kable(assoc_qt_params$data$ped_table)
```

```{r}
knitr::kable(assoc_qt_params$phe_table)
```

Should be equal:

```{r}
testthat::expect_equal(
  assoc_qt_params$data$map_table,
  read_plink_map_file(get_plinkr_filename("demo_assoc_qt.map"))
)
```

```{r}
testthat::expect_equal(
  assoc_qt_params$data$ped_table,
  read_plink_ped_file(get_plinkr_filename("demo_assoc_qt.ped"))
)
```

```{r}
if (1 == 2) {
  plinkr::save_phe_table(
    phe_table = assoc_qt_params$phe_table,
    phe_filename = get_plinkr_filename("demo_assoc_qt.phe")
  )
  testthat::expect_equal(
    assoc_qt_params$phe_table,
    read_plink_phe_file(get_plinkr_filename("demo_assoc_qt.phe"))
  )
}
```

### binary files

```{r}
if (1 == 2) {
  make_bed(
    base_input_filename = file.path(
      system.file("extdata", package = "plinkr"), "demo_assoc_qt"
    ),
    base_output_filename = file.path(
      system.file("extdata", package = "plinkr"), "demo_assoc_qt"
    )
  )

  make_bed(
    base_input_filename = file.path(
      system.file("extdata", package = "plinkr"), "demo_assoc"
    ),
    base_output_filename = file.path(
      system.file("extdata", package = "plinkr"), "demo_assoc"
    )
  )
}
```

# Cleanup

```{r cleanup}
unlink(folder_name)
```
